Tina Paper In Development

### 2024-10-31 docs/writing/outline.tex
I think I figured out the intro of the outline now. Writing in the paper is almost caught up
Starting to make methods more clear to read. Suddenly Latex workshop doesn't want to recognize references.

\1 Abstract
\begin{enumerate}
	\item Flatbed scanner alternative with improved resolution to digitizing tree cookies and cores
	\item Digital twins of samples is important for building data repositories and promoting replicable science
	\item Digitizing wood samples has limits to sample dimensions or lacks affordable implementations
	\item Current alternatives are very expensive (\$70,000USD), lack a high level of detail (printer scanners), or cannot scan cookies (CaptuRING)
	\item Tina is open source and able to be assembled with 3D printed parts, OpenBuilds parts kits, and hand tools % Can we add how much ours costs?
\end{enumerate}
\item Introduction
	\begin{enumerate}
	\item Tree ring science overview
		\begin{enumerate}
		\item tree rings provide critical data/insights across multiple fields % Guibal and Guiot
		\item tree rings can help reconstruct past climates (dendroclimatology) % Sheppard
		\item understanding tree growth mechanisms (plant bio/ecology) % Fritts and Swetnam 
		\end{enumerate}
	% Consider adding this to below points (Development of scanning alternatives) as one paragraph on its own
	\item Getting these data means measuring tree rings -- quick history
		\begin{enumerate}
		\item originally done on scopes (stage micrometer) % Robinson and Evans, https://conservancy.umn.edu/server/api/core/bitstreams/30f91f03-c27b-44de-8659-5156220ce79b/content
		\item eventually there was a shift to analyzing digital images of samples instead of stage micrometers % Guay et al. 1992 
		\item digital copies inherently allow for contributing to data pools  % Rademacher​ et al. WIAD; also International Tree-Ring Data Bank (IRTB) https://www.ncei.noaa.gov/products/paleoclimatology/tree-ring
		\item Development of image acquisition in dendrochronology 
			\begin{enumerate}
			\item two models: flatbed scanner or overlapping images and stitch
			\item beginning with flatbed scanners % WinDendro
				\subitem modern flatbed scanners can achieve high resolution (4200 dpi) but size of samples is limited % see CaptuRING paper for flatbed specs
			\item overlapping images and stitching 
				\subitem ATRICS, Gigapixel, CaptuRING 
				\subitem takes advantage of advanced camera technology
			\end{enumerate}
		% \item Could add here one sentence on the three methods you mention below depending on what you want to say ... 
		\end{enumerate}

@@

		\item Guiding design considerations
			\item Digitization of a tree cookie and core
			\item Minimizing barriers to assemble (Cost effective, open source, open hardware, minimal use of specialized equipment/tools)
			\item Reducing user interaction between sample prep and the final image
			\item Achieving near microscopic detail with a field of view as large as the sample

>>>>

\1 Abstract
\begin{enumerate}
	\item Flatbed scanner alternative with improved resolution to digitizing tree cookies and cores
	\item Digital twins of samples is important for building data repositories and promoting replicable science
	\item Digitizing wood samples has limits to sample dimensions or lacks affordable implementations
	\item Current alternatives are very expensive (\$70,000USD), lack a high level of detail (printer scanners), or cannot scan cookies (CaptuRING)
	\item Tina is open source and able to be assembled with 3D printed parts, OpenBuilds parts kits, and hand tools % Can we add how much ours costs?
\end{enumerate}
\item Introduction
	\begin{enumerate}
	\item Tree ring science overview
		\begin{enumerate}
		\item tree rings provide critical data/insights across multiple fields % Guibal and Guiot
		\item tree rings can help reconstruct past climates (dendroclimatology) % Sheppard
		\item understanding tree growth mechanisms (plant bio/ecology) % Fritts and Swetnam 
		\end{enumerate}
	% Consider adding this to below points (Development of scanning alternatives) as one paragraph on its own
	\item Getting these data means measuring tree rings -- quick history
		\begin{enumerate}
		\item originally done on scopes (stage micrometer) % Robinson and Evans, https://conservancy.umn.edu/server/api/core/bitstreams/30f91f03-c27b-44de-8659-5156220ce79b/content
		\item eventually there was a shift to analyzing digital images of samples instead of stage micrometers % Guay et al. 1992 
		\item digital copies inherently allow for contributing to data pools  % Rademacher​ et al. WIAD; also International Tree-Ring Data Bank (IRTB) https://www.ncei.noaa.gov/products/paleoclimatology/tree-ring
		\item Development of image acquisition in dendrochronology 
			\begin{enumerate}
			\item two models: flatbed scanner or overlapping images and stitch
			\item beginning with flatbed scanners % WinDendro
				\subitem modern flatbed scanners can achieve high resolution (4200 dpi) but size of samples is limited % see CaptuRING paper for flatbed specs
				\subitem can scan both cookies and cores	
			\item overlapping images and stitching 
				\subitem ATRICS, Gigapixel, CaptuRING 
				\subitem so far can only capture cores
				\subitem budget friendly implementations require user intervention between image acquisition and stitching
				\subitem takes advantage of advanced camera technology
			\end{enumerate}
		% \item Could add here one sentence on the three methods you mention below depending on what you want to say ... 
		\end{enumerate}

@@

		% 
		% \item Guiding design considerations
		% 	\begin{enumerate}
		% 	\item Digitization of a tree cookie and core
		% 	\item Minimizing barriers to assemble (Cost effective, open source, open hardware, minimal use of specialized equipment/tools)
		% 	\item Reducing user interaction between sample prep and the final image
		% 	\item Achieving near microscopic detail with a field of view as large as the sample
		% 	\end{enumerate}

### 2024-10-31 docs/writing/paper.tex
Starting to make methods more clear to read. Suddenly Latex workshop doesn't want to recognize references.

@@

$$\frac{1}{MN\mu} \sum\limits_{x=0}^{M-1} \sum\limits_{y=0}^{N-1}(f(x,y) - \mu)^2$$

\subsubsection{Automatic Control}
While the image stack provides a basis to obtain an in focus image, it is statically limited to the 1mm range in the Z axis. Instead of spending significant time guaranteeing that the sample
is within the the Z axis range across the entire surface area, a PID control algorithm takes the wheel \citep{odwyer_summary_2000}. The information previously calculated in the image stack is sufficient 
to inform the PID controller of how to adjust the initial the Z axis for the next image stack. The controller is constantly trying to make the most in focus image be at the center of the image stack,
index value 5. This allows for the most flexibility moving to the next location of the image grid.

\begin{figure}
    \centering
    \begin{subfigure}{.5\textwidth}
      \centering
      \includegraphics[height=0.75\linewidth]{../diagrams/sample_setup_ideal.png}
      \caption{Ideal sample leveling}
      \label{fig:sub1}
    \end{subfigure}%
    \begin{subfigure}{.5\textwidth}
      \centering
      \includegraphics[height=0.75\linewidth]{../diagrams/sample_setup_realistic.png}
      \caption{Realistic sample leveling}
      \label{fig:sub2}
    \end{subfigure}
    \caption{Side view of the camera and sample on top of a leveling table. The ideal sample leveling shows a uniform distance d at all (x,y) coordinates on the sample. This is impossible to achieve in reality, the true sample leveling has a non uniform distance at unique (x,y) coordinates.}
    \label{fig:test}
\end{figure}

>>>>

The frame, camera, computer, software, and sample leveling tables, work cohesively to address the nuances of digitizing both cookies and cores. 
Each of these subsystems will be reviewed along with the design considerations involved to make a complete machine. 
Considering Tina is an imaging machine, we will start off by looking at the camera and move outward.

\subsection{System Overview}
Camera, computer, software, XYZ frame, adapters.

\subsection{Camera}
In the Gigapixel and CaptuRING, a professional handheld camera was used to capture images. While this benefits from a wide variety of lens and camera choices - and principally autofocus - they come at a cost. 
Beyond its literal price tag, a professional camera also carries a significant weight, demanding a more robust and therefore more costly machine frame. 
These cameras are also optimized for handheld use, not for integration into computer driven machines. This results in clunky access to images from software. 
Designing around a machine vision camera such as the 12MP Raspberry Pi HQ Camera drastically reduced the camera cost while improving the access to the data stream coming from the camera sensor.
Pairing the Raspberry Pi HQ Camera with a high magnification C-Mount lens allows for images with very high resolution in dots-per-inch (DPI) to be captured. (THIS IS AN ESTIMATE, GET THE REAL VALUES FROM TINA) If the field of view of the image
could be optically zoomed to 5mm x 3mm, the resolution can reach a theoretical 32,000 DPI.

\subsubsection{Focusing}
Normalized variance process goes here.

Achieving a focused image across the entire surface area of the sample is the basis for an accurate stitch and effective analysis. By saving costs on a fixed focus camera and lens, it was necessary to implement a method to focus the images. 
Focus is solely a function of distance to the subject in the regime. With this high magnification lens, having an image go from in-focus to out-of-focus was sensitive to sub millimeter variation in distance to the lens. 
This requires cores and especially cookies to be as close to orthogonally aligned to the lens as possible. A 3D printed levelling table and bullseye level were used to correct plane misalignment of up to 10 degrees in cookies (CALCULATE THIS?). 
Slight sample misalignments are corrected within a closed loop automatic control PID algorithm.

@@

% $$\frac{1}{MN\mu} \sum\limits_{x=0}^{M-1} \sum\limits_{y=0}^{N-1}(f(x,y) - \mu)^2$$

% \subsubsection{Automatic Control}
% While the image stack provides a basis to obtain an in focus image, it is statically limited to the 1mm range in the Z axis. Instead of spending significant time guaranteeing that the sample
% is within the the Z axis range across the entire surface area, a PID control algorithm takes the wheel \citep{odwyer_summary_2000}. The information previously calculated in the image stack is sufficient 
% to inform the PID controller of how to adjust the initial the Z axis for the next image stack. The controller is constantly trying to make the most in focus image be at the center of the image stack,
% index value 5. This allows for the most flexibility moving to the next location of the image grid.

% \begin{figure}
%     \centering
%     \begin{subfigure}{.5\textwidth}
%       \centering
%       \includegraphics[height=0.75\linewidth]{../diagrams/sample_setup_ideal.png}
%       \caption{Ideal sample leveling}
%       \label{fig:sub1}
%     \end{subfigure}%
%     \begin{subfigure}{.5\textwidth}
%       \centering
%       \includegraphics[height=0.75\linewidth]{../diagrams/sample_setup_realistic.png}
%       \caption{Realistic sample leveling}
%       \label{fig:sub2}
%     \end{subfigure}
%     \caption{Side view of the camera and sample on top of a leveling table. The ideal sample leveling shows a uniform distance d at all (x,y) coordinates on the sample. This is impossible to achieve in reality, the true sample leveling has a non uniform distance at unique (x,y) coordinates.}
%     \label{fig:test}
% \end{figure}
